### 2018-11-15 Manuscript/Manuscript.tex
completion of implementation of round 1 edits

To assess the statistical significance of the correlation (or lack thereof) implied by the Pearson correlation coefficients, we simulated the four glacial episodes (Fig. \ref{fig:LIP_area}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and percent overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations (Fig. \ref{fig:LIP_correlation}). This approach accounts for the reality that spurious correlation can arise between auto-correlated data sets such as these, where each value is not independent, but is instead dependent on the previous state of the system. For the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario, 28\% of the randomly timed glacial intervals correlate better with the LIP area in the tropics. With an associated p-value of 0.28, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected. Taking this approach, none of the positive or negative correlations that emerge between the LIP area scenarios are statistically significant.

@@

We have focused this analysis on the Phanerozoic record given that well-constrained paleogeographic models are available for the past $\sim$520~Myr. Furthermore, the approach of seeking to evaluate correlation between LIP area and glaciation is complicated for Neoproterozoic Snowball Earth events, since ice-albedo runaway leads to persistent global glaciation absent of continued forcing through normal carbon cycle processes until sufficient CO$_2$ builds up in the absence of silicate weathering to drive deglaciation. Nevertheless, the hypothesis of tropical LIP area associated with the ca. 720~Ma Franklin LIP increasing global weatherability and contributing to the onset of the Sturtian Snowball Earth is a major motivating driver behind conducting this analysis. How does the tropical LIP area associated with the Franklin LIP compare to that observed in the Phanerozoic? Using the paleomagnetic pole of \citet{Denyszyn2009a}, we reconstruct the paleolatitude of the Franklin LIP at the time of emplacement, and find that $\sim$99.7\% (or $\sim$2.6~Mkm$^{2}$) of the LIP erupted within 15\textdegree\xspace of the equator. This Franklin LIP tropical area at the time of emplacement is approximately equivalent to the Cenozoic peak, and is smaller than the other Phanerozoic peaks (Fig. \ref{fig:LIP_area}C). The ca. 1109~Ma Umkondo is another Precambrian LIP that is constrained to have erupted in the tropics, although it is not known to be associated with any glaciation (no glacial deposits are found within the contemporaneous Midcontinent Rift basin; \citealp{Swanson-Hysell2018a}). We reconstruct the paleolatitude of the Umkondo LIP at the time of emplacement using the paleomagnetic pole of \citet{Swanson-Hysell2015b}, and find that effectively all of the LIP (or $\sim$2.0~Mkm$^{2}$) erupted within the tropics, an area that is slightly smaller than that for the Franklin LIP (Fig. \ref{fig:LIP_area}C). Together, these results indicate that the tropical area of the Franklin LIP, when compared to that of Phanerozoic as well as other Precambrian LIPs, did not have a singularly large area in the tropics. Given that similar peaks in tropical LIP area are not associated with the onset of glacial periods, additional processes must have been at play in drawing down atmospheric CO$_2$ in the lead-up to the Sturtian Snowball Earth. One such process could have been an increase in global weatherability associated with arc accretion within terranes of the present-day Arabian-Nubian Shield \citep{Park2018a}. An increase in global weatherability associated with the emplacement of the Franklin LIP in the tropics could have been a contributing factor in addition to other such processes.

The temporal overlap between Franklin LIP eruptions \citep{Denyszyn2009a} and the initiation of Sturtian glaciation \citep{Macdonald2010a, MacLennan2018a} remains compelling. This overlap continues to suggest that other aspects of LIP emplacement, such as the injection of sulfur aerosols in the stratosphere \citep{Macdonald2017a}, could have played a role in the initiation of low-latitude glaciation. The latitude of such aerosol injections is important for their temporary effect on albedo which is maximized at low latitude, and their presence at high concentrations is pre-conditioned on eruption through sedimentary basins hosting evaporite deposits \citep{Macdonald2017a}.

Overall, the conclusion that tropical LIP area associated with the Franklin is not uniquely high and therefore not a sole driver of Snowball Earth onset is consistent with the results of the Phanerozoic analysis. LIP area and its associated impact on global weatherability is not the fundamental control on Earth's long-term climate state.

>>>>

To assess the statistical significance of the correlation (or lack thereof) implied by the Pearson correlation coefficients, we simulated the four glacial episodes (Fig. \ref{fig:LIP_area}E) occurring at random times through the past 520~Myr, and recomputed the correlation coefficient and \% overlap between the LIP area in the tropics and the randomly timed glacial intervals for each of these 100,000 simulations (Fig. \ref{fig:LIP_correlation}). This approach accounts for the reality that spurious correlation can arise between auto-correlated data sets such as these, where each value is not independent, but is instead dependent on the previous state of the system. For the `t$_{1/2}$ = 120~Myr + 100\% burial' scenario, 28\% of the randomly timed glacial intervals correlate better with the LIP area in the tropics. With an associated p-value of 0.28, the null hypothesis that glacial intervals do not correlate to LIP area in the tropics cannot be rejected. Taking this approach, none of the positive or negative correlations that emerge between the LIP area scenarios are statistically significant.

@@

We have focused this analysis on the Phanerozoic record given that well-constrained paleogeographic models are available for the past $\sim$520~Myr. Furthermore, the approach of seeking to evaluate correlation between LIP area and glaciation is complicated for Neoproterozoic Snowball Earth events, since ice-albedo runaway leads to persistent global glaciation absent of continued forcing through normal carbon cycle processes until sufficient CO$_2$ builds up in the absence of silicate weathering to drive deglaciation. Nevertheless, the hypothesis of tropical LIP area associated with the ca. 720~Ma Franklin LIP increasing global weatherability and contributing to the onset of the Sturtian Snowball Earth is a major motivating driver behind conducting this analysis. How does the tropical LIP area associated with the Franklin LIP compare to that observed in the Phanerozoic? Using the paleomagnetic pole of \citet{Denyszyn2009a}, we reconstruct the paleolatitude of the Franklin LIP at the time of emplacement, and find that $\sim$99.7\% (or $\sim$2.6~Mkm$^{2}$) of the LIP erupted within 15\textdegree\xspace of the equator. This Franklin LIP tropical area at the time of emplacement is approximately equivalent to the Cenozoic peak, and is smaller than the other Phanerozoic peaks (Fig. \ref{fig:LIP_area}C). The ca. 1109~Ma Umkondo is another Precambrian LIP that is constrained to have erupted in the tropics, although it is not known to be associated with any glaciation (no glacial deposits are found within the contemporaneous Midcontinent Rift basin; \citealp{Swanson-Hysell2018a}). We reconstruct the paleolatitude of the Umkondo LIP at the time of emplacement using the paleomagnetic pole of \citet{Swanson-Hysell2015b}, and find that effectively all of the LIP (or $\sim$2.0~Mkm$^{2}$) erupted within the tropics, an area that is slightly smaller than that for the Franklin LIP (Fig. \ref{fig:LIP_area}C). Together, these results indicate that the tropical area of the Franklin LIP, when compared to that of Phanerozoic as well as other Precambrian LIPs, did not have a singularly large area in the tropics. Given that similar peaks in tropical LIP area are not associated with the onset of glacial periods, additional processes must have been at play in drawing down atmospheric CO$_2$ in the lead-up to the Sturtian Snowball Earth. One such process could have been an increase in global weatherability associated with arc accretion within terranes of the present-day Arabian-Nubian Shield \citep{Park2018a}. An increase in global weatherability associated with the emplacement of the Franklin LIP in the tropics could have been a contributing factor in addition to other such processes. The temporal overlap between Franklin LIP eruptions \citep{Denyszyn2009a} and the initiation of Sturtian glaciation \citep{Macdonald2010a, MacLennan2018a} remains compelling. This overlap continues to suggest that other aspects of LIP emplacement, such as the injection of sulfur aerosols in the stratosphere \citep{Macdonald2017a}, could have played a role in the initiation of low-latitude glaciation. The latitude of such aerosol injections is important for their temporary effect on albedo which is maximized at low latitude, and their presence at high concentrations is pre-conditioned on eruption through sedimentary basins hosting evaporite deposits \citep{Macdonald2017a}. Overall, the conclusion that tropical LIP area associated with the Franklin is not uniquely high and therefore not a sole driver of Snowball Earth onset is consistent with the results of the Phanerozoic analysis. LIP area and its associated impact on global weatherability is not the fundamental control on Earth's long-term climate state.
